Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
conclusions 3

Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\sim v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ may increase even more rapidly as $v_w$ decreases for two reasons: (1) when $v_w$ is comparable or less than the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy at infinity (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on previous studies (e.g.,\citealt{Ciotti+10}) is found to result in time-dependent cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$).

\item Galaxies with $10^{6}M_{\odot} \lesssim \Mbh \lesssim 10^8
  \Msun$ with star formation histories matching the average values
  derived from the cosmological simulations of
  \citet{MosterNaab+:2013a} possess an effective heating rate $v_w\sim
  1000$ km s$^{-1}$, relatively independent of SMBH mass
  (Fig.~\ref{fig:vwSources}).  This results in a prediction for the
  average Eddington ratio $\eddr \propto \Mbh^{\alpha}$ with
  $\alpha\lsim 0.5 \, (0.8)$ for core (cusp) galaxies which is tension
  with the trend of measured nuclear X-ray luminosities of early-type
  galaxies (e.g., \citealt{Miller+15}), which instead find that
  $\eddr$ increases towards smaller $\Mbh$.  However, high X-ray
  radiative efficiencies are required for these galaxies to be
  consistent with thermally-stable accretion, instead suggesting that
  (at least at low black hole masses), the possibility of two branches
  of accretion states.

    % However, it is important to note that the inflow rate on large
    % scales (which we calculate) may differ from
    % the actual accretion rate as outflows could occur on small
    % scales. The outflow rate would depend on the angular momentum of
    % the infalling gas and thus the angular momentum of stars in the
    % nucleus. The in turn would vary systematically with $\Mbh$.

    For $\Mbh\gsim 10^8 \Msun$, the flow from stellar winds onto the
    BH is more susceptible to thermal instability due to the lower
    effective wind velocity $v_w$. 
  \item We expect scatter in nuclear x-ray luminosities at
    fixed $\Mbh$ from (1) stochastic variation in star formation
    history from galaxy to galaxy (2) Variations in small scale outflow
    rates due to variations between galactic nuclear angular
    momenta. {\bf AG: some more explanation of why 1 might not show up
    in Pellegrini...}
    These variations alone could explain much of the 2-4 orders of
    magnitude of scatter of nuclear x-ray luminosity at fixed $\Mbh$ 
    in the samples of \citet{Miller+15} and \citet{Pellegrini:2010a}.

    It is also possible that many of the galaxies in the
    \citet{Miller+15} are accreting in a non-steady fashion and this
    scatter arises from cyclic variation in accretion rate. In fact it
    that a population of cyclically accreting with $\Mbh\gsim 10^{8}
    \Msun$ coexists with a population of smaller, steadily accreting
    black holes (e.g. the upper limits in \citet{Miller+15}).  This
    calls into question the use of simple extrapolations of power-law
    fits to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$
    to constrain the occupation fraction of black holes in low mass
    galaxies (\citealt{Miller+15}).
  \item Low mass AGN have growth times comparable to the Hubble time
    (\citet{Heckman+04}). We have shown that this would require accretion
    rates not consistent with steady-state thermally stable accretion
    of stellar wind material, and is driven by an external supply of gas.
  \item We compare the $\Mdot$ for each of our solution to the one
    calculated from the Bondi formula. We find that they agree to
    within a factor of order unity. This is because in our model the
    accretion rate is determined by the flow properties at the
    stagnation radius, $\rs$, which is within a factor of a few of the
    Bondi radius $\rb$. Thus, use of the Bondi formula is a
    good approximation.
    \end{enumerate}

>>>>

Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\sim v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ may increase even more rapidly as $v_w$ decreases for two reasons: (1) when $v_w$ is comparable or less than the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy interior to the break radius of the stellar density profile (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on previous studies (e.g.,\citealt{Ciotti+10}) is found to result in time-dependent cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$).

\item The nuclei of galaxies with low mass black holes $\Mbh \lesssim 10^8
  \Msun$ and star formation histories which match their average value derived from cosmological simulations, receive a stellar wind heating rate $v_w\sim 800$ km s$^{-1}$ which is relatively independent of SMBH mass (Fig.~\ref{fig:vwSources}).  The resulting Eddington ratio is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in moderate tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).

For thermally stable accretion models, scatter in nuclear X-ray luminosities at fixed $\Mbh$ of the observed 2-4 orders of magnitude can result from a combination of (1) intrinsic scatter in the SMBH scaling relations; (2) stochastic variation in star formation history from galaxy to galaxy, or from time to time due to burstiness in the star formation history (e.g.~\ref{fig:NickPlot2}); (3) Variations in the amount of inflowing mass which is lost to small scale outflows from the SMBH accretion disk, due to variations in the angular momentum of the accreting gas.

\item Only for relatively high X-ray radiative efficiencies similar to the maximum values predicted by ADAF models are the inferred accretion rates of quiescent early-type galaxies below the maximum value consistent with thermally-stable accretion (Fig.~\ref{fig:miller}).  High mass SMBHs are particularly susceptible to thermal instability, given the large stagnation radii that can result from the deep potentials and large break radii of these (mainly core) galaxies.  If the current X-ray sample is instead comprised mainly of thermally unstable nuclei, then a separate branch of low $L_x$ "thermally stable" nuclei may exist, especially at low $M_{\bullet}$.  Such a dichotomy calls into question the accuracy of extrapolating the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ in constraining the SMBH occupation fraction in low mass galaxies (\citealt{Miller+15}).

  \item Low mass AGN have growth times comparable to the Hubble time (\citet{Heckman+04}).  Our results show that this would require accretion rates which are not consistent with steady-state, thermally stable accretion of stellar wind material.  Low mass black hole growth must instead be driven by an external supply of gas.
  \item We compare the $\Mdot$ for each of our solution to the one
    calculated from the Bondi formula. We find that they agree to
    within a factor of order unity. This is because in our model the
    accretion rate is determined by the flow properties at the
    stagnation radius, $\rs$, which is within a factor of a few of the
    Bondi radius $\rb$. Thus, use of the Bondi formula is a
    good approximation.
    \end{enumerate}
